---
title: "Marketing Strategy"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Problem Definition
#### 1. Specifying Analysis Question
Perform cluster analysis on customer behaviour data collected by Kira Plastinina, a russian brand to help the sales team identify characteristics of customer groups.

#### 2. Metrics of success
Specification of distinct customer groups and their distinctive characteristics.

#### 3. Understanding the context
Kira Plastinina is a Russian brand that is sold through a defunct chain of retail stores in Russia, Ukraine, Kazakhstan, Belarus, China, Philippines, and Armenia. The brand’s Sales and Marketing team would like to understand their customer’s behavior from data that they have collected over the past year. More specifically, they would like to learn the characteristics of customer groups.

#### 4. Recording the experimental design
- Data Sourcing
- Check the Data
- Perform Data Cleaning
- Perform Exploratory Data Analysis  (Univariate, Bivariate & Multivariate)
- Implement the Solution
- Challenge the Solution
- Follow up Questions

## Data Sourcing
The dataset for this project is provided as part of the Moringa curriculum. It consists of 18 attributes, 10 numerical and 8 categorical. It consists of metadata collected from customers that visit the Kira Plastinina e-commerce website over the period of one year.

## Data Loading and Inspection
```{r}
# Load libraries
library(readr)
library(tidyverse)
library(psych)
library(corrplot)
library(BBmisc)
library(caret)

```
```{r}
shoppers <- read_csv("shoppers.csv")

# Preview data
head(shoppers)
```
```{r}
# Shape of data
dim(shoppers);

# Column datatypes
str(shoppers);
```


```{r}
# Get summary of data
summary(shoppers)
```
A number of columns contain null values: ProductRelated,ProductRelated_Duration,BounceRates,ExitRates, Administrative, Administrative_Duration,Informational,Informational_Duration

There appears to be anomalies in the ProductRelated_Duration, Administrative_Duration and Informational_Duration columns with certain examples having a value of -1

```{r}
# Check for duplicated data
anyDuplicated(shoppers)
```
159 duplicated rows in the data.

```{r}
# Check for unique values in month and visitor type columns
unique(shoppers$Month);
unique(shoppers$VisitorType);
```

##  Data Cleaning

```{r}
# Get number of rows with null values
shoppers[!complete.cases(shoppers),]

```
All fourteen rows have missing values in all the following columns: ProductRelated, ProductRelated_Duration, BounceRates, ExitRates, Administrative, Administrative_Duration, Informational and Informational_Duration. Drop rows.

```{r}
# Drop rows
shoppers <- shoppers[complete.cases(shoppers),]

# Drop duplicated rows
shoppers <- shoppers[!duplicated(shoppers), ]
```

```{r}
# Investigate rows with -1 in duration columns
anomaly <- shoppers %>% select(c(Administrative_Duration, Administrative, Informational_Duration, Informational, ProductRelated_Duration, ProductRelated)) %>% filter(Administrative_Duration == -1 | Informational_Duration == -1 | ProductRelated_Duration == -1)

anomaly
```
33 rows have values of -1 in all three duration columns, which is an anomaly in the data as it means that the customer did not spend any time on any page category be it administrative, informational or product related.

```{r}
# Drop anomalous data
shoppers <- shoppers %>% filter(Administrative_Duration != -1, Informational_Duration != -1, ProductRelated_Duration != 1)
```


```{r}
# describe each column
describe(shoppers[, c(1:9)])
```
All variables are heavily positively skewed and have high kurtosis values with Informational_Duration, ProductRelated_Duration and PageValues being the most heavily skewed. 
```{r}
# Custom function for number of outliers per column
outlier_detector <- function(x){
  out <- boxplot.stats(x)$out
  return((length(out)/  12164)*100)
}
  
# Get outlier count per column
sapply(shoppers[,c(1:9)], outlier_detector)
```
```{r}
# Plot boxplots of columns with high % of outliers

boxplot(shoppers$Informational, shoppers$Informational_Duration, shoppers$PageValues,
main = "Columns with high values of outliers",
names = c("Informational", "Informational_Duration", "Page Values"),
col = c("orange","red"),
border = "brown",
notch = TRUE)
```

All outliers are contained above the third quantile, this indicates that all outliers are contained in the higher ranges of the above variables. Considering the nature of the data, it is entirely possible for customers to spend long durations on informational pages or be browsing pages that have items with high values.Therefore maintain outliers.

## Exploratory Data Analysis
```{r}
# Frequency distribution of categorical columns
sapply(shoppers[, c(10:18)], table)
```

*Summary of frequency table*

- Months with the highest traffic are May, November, March and December.
- Most visitors have a type 2 operating system.
- Most visitors have a type 2 browser.
- Most visitors to the site are located in region 1
- Most of the traffic to the website is of type 2 and 1
- Visitors to the site are mostly returning visitors.
- Most visits to the site do not earn revenue

```{r}
# Number of visits to product related pages per month
product_stats <- shoppers %>% select(ProductRelated, ProductRelated_Duration, Month)%>%group_by(Month)%>% summarise_all(mean)

product_stats[order(product_stats$ProductRelated, decreasing = TRUE),]
```

November had the highest number of visits to product related pages with the highest duration. 

```{r}
# Get the distribution of revenue generating visits among visitors
shoppers %>% select(VisitorType, Revenue)%>% group_by(VisitorType, Revenue)%>% summarise(visits = n())%>% filter(Revenue==TRUE)
```

Returning visitors have the highest number of revenue generating visitors.

```{r}
# Get the distribution of revenue generating visits among regions
shoppers %>% select(Region, Revenue)%>% group_by(Region, Revenue)%>% summarise(visits = n())%>% filter(Revenue==TRUE)
```
Visitors located in region 1 have the highest number of revenue generating visits.

```{r}
# Get page values and visits to product pages based on proximity to a special day
special_product <- shoppers %>% select(SpecialDay, PageValues, ProductRelated, ProductRelated_Duration)%>% group_by(SpecialDay)%>% summarise_all(mean)


ggplot(special_product, aes(x = factor(SpecialDay), y=ProductRelated))+
  geom_bar(fill = 'lightblue', stat = "identity", width = 0.6) +
  labs(x = 'SpecialDay')+
  coord_flip()
```

On special days customers visit the highest number of product related pages, with the highest mean number of pages visited.

```{r}
# Distribution of revenue generating visits based on proximity to special days
revenue_special <- shoppers %>% select(SpecialDay, Revenue)%>% group_by(SpecialDay, Revenue)%>% summarise(count = n())%>% mutate(freq = count / sum(count)) %>% filter(Revenue==TRUE)%>% mutate(freq = round((freq * 100), 1)) 

revenue_special <- revenue_special[order(revenue_special$freq, decreasing = TRUE),]

ggplot(revenue_special, aes(x = factor(SpecialDay), y=freq))+
  geom_bar(fill = 'lightblue', stat = "identity", width = 0.6) +
  labs(y = '% of revenue generating', x = 'Special Days' )+
  coord_flip()

```


Ordinary days have the highest percentage of visits that are income generating followed by days 0.6 and 0.2 away from special days.

```{r}
# Get bounce rates and exit rates among visitor groups
visitor_stats <- shoppers %>% select(VisitorType, ExitRates, BounceRates)%>% group_by(VisitorType)%>%summarise_all(mean) 

visitor_stats

```

Visitors of type other have the highest exit rates and bounce rates followed by returning visitors.

```{r}
# Page views and durations based on visitor type
shoppers %>% select(VisitorType, Administrative:ProductRelated_Duration)%>% group_by(VisitorType)%>%summarise_all(mean)

```

New visitors have the highest mean number of administrative pages visited as well as the longest administrative duration. Returning visitors mostly visit informational and product related pages and spend the longer durations on these pages than other visitor types.

```{r}
traffic_stats <- shoppers %>% select(TrafficType, ExitRates, BounceRates)%>% group_by(TrafficType)%>% summarise_all(mean)

par(mfrow = c(1,2))

ggplot(traffic_stats, aes(x=TrafficType, y = ExitRates))+
  geom_bar(stat = "identity", fill="peachpuff2")

ggplot(traffic_stats, aes(x=TrafficType, y = BounceRates))+
  geom_bar(stat = "identity", fill="peachpuff2")


```

Traffic of type 15 and 17 have the highest mean exit rates and bounce rates.

```{r}
traffic_page_stats <- shoppers %>% select(TrafficType, Administrative:ProductRelated_Duration)%>% group_by(TrafficType)%>% summarise_all(mean)

par(mfrow = c(1,3))

ggplot(traffic_page_stats, aes(x=TrafficType, y = Administrative))+
  geom_bar(stat = "identity", fill="lightblue")

ggplot(traffic_page_stats, aes(x=TrafficType, y = Informational))+
  geom_bar(stat = "identity", fill="lightblue")

ggplot(traffic_page_stats, aes(x=TrafficType, y = ProductRelated))+
  geom_bar(stat = "identity", fill="lightblue")
```

Traffic type 14 has the highest mean number of product related and infomational pages visited. Traffic type 4 and 14 have the highest mean number of administrative pages visited.

```{r}

corrplot(corr = cor(shoppers[, c(1:9)]), method = "number", type = "upper", order = "hclust", tl.col = "black", tl.cex = 0.6)
```

As expected, there is a high correlation between each type page and the corresponding page duration. There is also a high correlation between bounce rates and exit rates.

## Solution Implementation
### K-Means Clustering
```{r}
# One hot encode categorical features
new_shoppers <- shoppers[, 1:17]
new_shoppers[, 10:16] <- as.character(new_shoppers[, 10:16])
dmy <- dummyVars("~ SpecialDay + Month + OperatingSystems + Browser + Region + TrafficType + VisitorType", data= new_shoppers)

ohe <- data.frame(predict(dmy, newdata = new_shoppers))

model_data <- cbind(new_shoppers[ , 1:9], ohe)

model_data$Weekend <- as.numeric(shoppers$Weekend)

head(model_data)
```
```{r}
# Normalize continous variables
model_data[, 1:9] <- normalize(model_data[, 1:9], method = "range")
```
```{r}
library(cluster)
library(factoextra)
library(NbClust)

# Perform kmean clustering with 5 clusters
customer_groups <- kmeans(model_data, centers = 5, nstart = 25) 

str(customer_groups)
```

```{r}
# Size of each cluster
customer_groups$size
```
The first cluster has the most components
```{r}
# Visualize  clusters

fviz_cluster(customer_groups, data = model_data)
```

```{r}
# The visualization above shows the data is better fitted to 7 clusters
customer_groups <- kmeans(model_data, centers = 7, nstart = 25)

customer_groups$size

```
```{r}
fviz_cluster(customer_groups, data = model_data)

```
After dimension reduction using PCA, the data has been separated into 7 distinct clusters.

### Hierachical Clustering
```{r}
# Determine best value for epislon
dbscan::kNNdistplot(model_data, k =  7)
abline(h = 0.23, lty = 2)
```


```{r}
db_groups <- dbscan::dbscan(model_data, eps = 0.23, minPts = 7)

db_groups
```

```{r}
fviz_cluster(db_groups, model_data, geom="point")
```
 
 The clusters in our data are spherical, thus the data naturally clusters into its seven distinct clusters despite dbscan identifying 23clusters.
 
 